Red-Giant Hot Jupiters: Brilliant Radio Beacons?

### 2015-09-06 ms.tex
Move comment to new line.
Figure 3 is updated

\author{
%
Yuka Fujii\altaffilmark{0} \\
%
David S. Spiegel\altaffilmark{1,2} \\
%
{\bf and some order:} \\
%
Jason Nordhaus\altaffilmark{3,4} \\
%
Nikku Madhusudhan\altaffilmark{5} \\
%
Mehrdad Mirbabayi\altaffilmark{1} \\
%
Aaron Parsons\altaffilmark{6} \\
%
Tony Mroczkowski\altaffilmark{7,8} \\
%
Neil T. Zimmerman\altaffilmark{9}
}

\affil{$^0$Earth-Life Science Institute, Tokyo Institute of Technology, 
  Tokyo, JAPAN, 152-8550}

\affil{$^1$Astrophysics Department, Institute for Advanced Study,
  Princeton, NJ 08540}

\affil{$^2$Research \& Development, Sum Labs,
  New York, NY  10001}

\affil{$^3$Department of Science and Mathematics, National Technical Institute for the Deaf, Rochester Institute of Technology, Rochester, NY 14623}

\affil{$^4$Center for Computational Relativity and Gravitation, Department of Mathematics, Rochester Institute of Technology, Rochester, NY 14623}

\affil{$^5$Astronomy Department, University of Cambridge, UK}

\affil{$^6$Astronomy Department, University of California Berkeley}

\affil{$^7$National Research Council Fellow}

\affil{$^8$Naval Research Laboratory, 4555 Overlook Ave SW, Washington, DC 20375}

\affil{$^9$Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218}

@@

Based on equations (\ref{eq:scale_Mdot}) and (\ref{eq:scale_v}), the number density of the stellar wind (equation (\ref{eq:n})) are normalized as follows:
%%%%%%%%%% 
\begin{eqnarray}
n &=& 1.80 \times 10^6 \; {\rm cm^{-3}} \\
&&\times \left( \frac{\dot M_\star}{10^{-8} M_{\odot }{\rm /yr}} \right) \left( \frac{v}{30~\mbox{km/s}} \right)^{-1} \left( \frac{a}{5 \; {\rm AU}} \right)^{-2}. \label{eq:n_normalized}
\end{eqnarray}
%%%%%%%%%%

As discussed in \citet{ignace2010}, the ionization fraction of the stellar wind of evolved stars are supposedly as low as $\sim 10^{-3}$\memoYF{cite something}. Since only charged particles interact with planetary magnetic field, this suggests the inefficient interaction with planetary magnetosphere and hence low input energy for radio emission. 
However, at this stage, the velocity of stellar wind becomes slower than the escape velocity of planetary companion and hence the stellar wind particles are expected to accrete onto the planets. 
\citet{spiegel+madhusudhan2012} computed the accretion luminosity $L_{\rm acc}$ and the temperature $T_{\rm acc}$ are: 
%%%%%%%%%
\begin{eqnarray}
L_{\rm acc} \sim &&  10^{28} \; {\rm erg\;s}^{-1} \left( \frac{\dot M_*}{10^{-8} M_{\odot }/{\rm yr}} \right)  \\
&& \times \left( \frac{M_p}{10 M_J} \right)^3 \left( \frac{M_*}{M_{\odot }} \right) \\
%
T_{\rm acc} \sim && 2 \times 10^6 ~\mbox{K} \left( \frac{M_p}{10 M_J} \right) \\
\sim && 2 \times 10^5 ~\mbox{K} \left( \frac{M_p}{M_J} \right). 
\end{eqnarray}
%%%%%%%%%
The accretion disks around planets therefore emit X-ray radiation. 
We estimate the scale of the ionized region by computing Stromgren radius of the accretion disk \citep{stromgren1939}: 
%%%%%%%%%
\begin{equation}
R_{\rm stromgren} = \frac{3}{4\pi} \left( \frac{S}{n^2 \beta } \right)^{1/3}
\end{equation}
%%%%%%%%%
where $S$ is the source rate of ionizing photons, $n$ is the density of hydrogen, and $\beta $ is the coefficient for recombination \memoYF{cite something}:
% http://www.tapir.caltech.edu/~chirata/ay102/Photoionized.pdf
% T \propto 4.3e-13 ( T /10^4 K )^-0.7
%%%%%%%%%
\begin{eqnarray}
%\beta &=& 4.3? \times 10^{-13} \; \left( \frac{T}{10^4 \;{\rm K}} \right) ^{-3/4} \;\; {\rm cm^3 \; sec^{-1}} \\
\beta &=& 2.6 \times 10^{-13} \; \left( \frac{T}{10^4 \;{\rm K}} \right) ^{-3/4} \;\; {\rm cm^3 \; sec^{-1}} \\
&\sim& 2.74 \times 10^{-14} \; \left( \frac{M_p}{M_J} \right)^{-3/4} \;\; {\rm cm^3 \; sec^{-1}} 
\end{eqnarray}
%%%%%%%%%
The source rate is obtained by counting the number of photons whose energy is higher than 13.6~eV ($\lambda = $), the energy needed to ionize hydrogen atoms.
As a result, 
%%%%%%%%%
\begin{eqnarray}
S_* \sim  
  \left\{
    \begin{array}{ll}
      2 \times 10^{42} \;  {\rm sec^{-1}} & \;\;\;(M_p = M_J) \\
      3 \times 10^{44} \; {\rm sec^{-1}} & \;\;\;(M_p = 10 M_J)
    \end{array}. 
  \right.
\end{eqnarray}
%%%%%%%%%
Using the number density of hydrogen atoms given by equation (\ref{eq:n_normalized}), 
%%%%%%%%%
\begin{eqnarray}
R_{\rm stromgren} \sim  
  \left\{
    \begin{array}{ll}
      4 \;{\rm AU} & \;\;\;(M_p = M_J) \\
      40 \;{\rm AU} & \;\;\;(M_p = 10 M_J)
    \end{array}
  \right.
\end{eqnarray}
%%%%%%%%%
for a planet around a red giants at 5~AU.

@@

Radio emission from the planet is observable from ground only when the maximum frequency is larger than the plasma frequency of Earth's ionosphere $f_{\rm plasma}^\oplus$ and the maximum plasma frequency along the line of sight $f_{\rm plasma}^{\rm los}$.
The plasma frequency may be expressed as
\begin{eqnarray}
f_{\rm plasma} & = & \sqrt{\frac{n_e e^2}{\pi m_e}} \\
 & = & 8979 {\rm~Hz} \times \left( \frac{n_e}{\rm cm^{-3}} \right)^{1/2} \, .
\label{eq:fplasma}
\end{eqnarray}
In the Earth's ionosphere, the electron number density is less than $10^6$~cm$^{-3}$, which implies that $f_{\rm plasma}^\oplus \lsim 10$~MHz.
At the favorable viewing condition, when the planet is on the near side of its star to the Earth, the maximum plasma frequency along the line of sight corresponds to that in the vicinity of the planet.
Therefore, combining Eq.~(\ref{eq:n}) and Eq.~(\ref{eq:fplasma}), we have
%%%
\begin{eqnarray}
\label{eq:fplasmalos} f_{\rm plasma}^{\rm los} & \sim & 8.98 {\rm~kHz} \times \left( \frac{\dot M_\star}{4 \pi a^2 m_p v} \times 1\rm~cm^3 \right)^{1/2} \\
\label{eq:fplasmalos_scaled} & = & 14.7{\rm~MHz} \times \left( \frac{a}{5~\mbox{AU}}\right)^{-1} \left(\frac{v}{30~\mbox{km/s}}  \right)^{-1/2} \notag \\
 & & \times \left( \frac{\dot M_\star}{10^{-8} M_{\odot }{\rm /yr}}\right)^{1/2} \, .
\end{eqnarray}
\memoYF{check numbers in above equation}
%%%
%where in Eq.~(\ref{eq:fplasmalos_scaled}) we have taken the solar mass-loss rate $\dot M_\odot$ to be $2.0 \times 10^{-14}~M_\odot$/yr and the typical solar wind velocity $v_{\odot }$ to be 400~km/sec.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[bp]
   \plotonesc{radio_emission_Mdot1e-8_constRp_christensen.pdf}
   \plotonesc{radio_emission_Mdot1e-5_constRp_christensen.pdf}
   \caption{[{\bf New version: based on scaling relationship of Christensen 2010}] Top panel: Intensity of radio emission (in the unit of Jy) from a companion to a RG star (left) and an AGB star at 10 pc.
The hashed region with vertical lines are not observable because of the plasma frequency cut-off of Earth's ionosphere. 
he hashed region with horizontal lines are not observable because of the plasma frequency cut-off of the stellar wind plasma in the vicinity of the companion.
Bottom panel: cyclotron frequency, i.e., the frequency of radio emission. }
  \label{fig:radio}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

One of the key features that make planetary radio emission distinct is the almost complete polarization \citep[e.g.][]{dulk1994}. ...\memoYF{write here}
s
Another key is that auroral radio emission from planets is not a static source but it significantly varies in time. 
The potential factors that cause time variability of the auroral radio emission are  listed below.

>>>>

\author{
%
Yuka Fujii\altaffilmark{1} 
%
David S. Spiegel\altaffilmark{2,3}
%
Tony Mroczkowski\altaffilmark{4,5}
%
Jason Nordhaus\altaffilmark{6,7}\\
%
Neil T. Zimmerman\altaffilmark{8}
%
Aaron Parsons\altaffilmark{9}
%
Nikku Madhusudhan\altaffilmark{10}
%
Mehrdad Mirbabayi\altaffilmark{2}
}

\affil{$^1$Earth-Life Science Institute, Tokyo Institute of Technology, 
  Tokyo, JAPAN, 152-8550}

\affil{$^2$Astrophysics Department, Institute for Advanced Study,
  Princeton, NJ 08540}

\affil{$^3$Research \& Development, Sum Labs,
  New York, NY  10001}

\affil{$^4$National Research Council Fellow}

\affil{$^5$Naval Research Laboratory, 4555 Overlook Ave SW, Washington, DC 20375}

\affil{$^6$Department of Science and Mathematics, National Technical Institute for the Deaf, Rochester Institute of Technology, Rochester, NY 14623}

\affil{$^7$Center for Computational Relativity and Gravitation, Department of Mathematics, Rochester Institute of Technology, Rochester, NY 14623}

\affil{$^8$Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218}

\affil{$^9$Astronomy Department, University of California Berkeley}

\affil{$^10$Astronomy Department, University of Cambridge, UK}

@@

Based on equations (\ref{eq:scale_Mdot}) and (\ref{eq:scale_v}), the number density of the stellar wind (equation (\ref{eq:n})) are normalized as follows:
%%%%%%%%%% 
\begin{eqnarray}
n &=& 1.80 \times 10^6 \; {\rm cm^{-3}} \\
&&\times \left( \frac{\dot M_\star}{10^{-8} M_{\odot }{\rm /yr}} \right) \left( \frac{v}{30~\mbox{km/s}} \right)^{-1} \left( \frac{a}{5 \; {\rm AU}} \right)^{-2}. \label{eq:n_normalized}
\end{eqnarray}
%%%%%%%%%% 
\memoYF{check the values above}

As discussed in \citet{ignace2010}, the ionization fraction of the stellar wind of evolved stars are supposedly as low as $\sim 10^{-3}$\memoYF{cite something}. Since only charged particles interact with planetary magnetic field, this suggests the inefficient interaction with planetary magnetosphere and hence low input energy for radio emission. 
However, at this stage, the velocity of stellar wind becomes slower than the escape velocity of planetary companion and hence the stellar wind particles are expected to accrete onto the planets. 
\citet{spiegel+madhusudhan2012} computed the accretion luminosity $L_{\rm acc}$ and the temperature $T_{\rm acc}$ are: 
%%%%%%%%%
\begin{eqnarray}
L_{\rm acc} \sim &&  10^{28} \; {\rm erg\;s}^{-1} \left( \frac{\dot M_*}{10^{-8} M_{\odot }/{\rm yr}} \right)  \notag \\
&& \times \left( \frac{M_p}{10 M_J} \right)^3 \left( \frac{M_*}{M_{\odot }} \right) \\
%
T_{\rm acc} \sim && 2 \times 10^6 ~\mbox{K} \left( \frac{M_p}{10 M_J} \right) \\
\sim && 2 \times 10^5 ~\mbox{K} \left( \frac{M_p}{M_J} \right). 
\end{eqnarray}
%%%%%%%%%
The accretion disks around planets therefore emit X-ray radiation. 
We estimate the scale of the ionized region by computing Stromgren radius of the accretion disk \citep{stromgren1939}: 
%%%%%%%%%
\begin{equation}
R_{\rm stromgren} = \frac{3}{4\pi} \left( \frac{S}{n^2 \beta } \right)^{1/3}
\end{equation}
%%%%%%%%%
where $S$ is the source rate of ionizing photons, $n$ is the density of hydrogen, and $\beta $ is the coefficient for recombination \memoYF{cite something}:
% http://www.tapir.caltech.edu/~chirata/ay102/Photoionized.pdf
% T \propto 4.3e-13 ( T /10^4 K )^-0.7
%%%%%%%%%
\begin{eqnarray}
%\beta &=& 4.3? \times 10^{-13} \; \left( \frac{T}{10^4 \;{\rm K}} \right) ^{-3/4} \;\; {\rm cm^3 \; sec^{-1}} \\
\beta &=& 2.6 \times 10^{-13} \; \left( \frac{T}{10^4 \;{\rm K}} \right) ^{-3/4} \;\; {\rm cm^3 \; sec^{-1}} \\
&\sim& 2.74 \times 10^{-14} \; \left( \frac{M_p}{M_J} \right)^{-3/4} \;\; {\rm cm^3 \; sec^{-1}} 
\end{eqnarray}
%%%%%%%%%
The source rate is obtained by counting the number of photons whose energy is higher than 13.6~eV ($\lambda = $), the energy needed to ionize hydrogen atoms.
As a result, 
%%%%%%%%%
\begin{eqnarray}
S_* \sim  
  \left\{
    \begin{array}{ll}
      2 \times 10^{42} \;  {\rm sec^{-1}} & \;\;\;(M_p = M_J) \\
      3 \times 10^{44} \; {\rm sec^{-1}} & \;\;\;(M_p = 10 M_J)
    \end{array}. 
  \right.
\end{eqnarray}
%%%%%%%%%
Using the number density of hydrogen atoms given by equation (\ref{eq:n_normalized}), 
%%%%%%%%%
\begin{eqnarray}
R_{\rm stromgren} \sim  
  \left\{
    \begin{array}{ll}
      4 \;{\rm AU} & \;\;\;(M_p = M_J) \\
      40 \;{\rm AU} & \;\;\;(M_p = 10 M_J)
    \end{array}
  \right.
\end{eqnarray}
%%%%%%%%%
for a planet around a red giants at 5~AU.

@@

Radio emission from the planet is observable from ground only when the frequencies of the radio emission are larger than the plasma frequency of Earth's ionosphere $f_{\rm plasma}^\oplus$ and the maximum plasma frequency along the line of sight $f_{\rm plasma}^{\rm los}$.
The plasma frequency may be expressed as
\begin{eqnarray}
f_{\rm plasma} & = & \sqrt{\frac{n_e e^2}{\pi m_e}} \\
 & = & 8979 {\rm~Hz} \times \left( \frac{n_e}{\rm cm^{-3}} \right)^{1/2} \, .
\label{eq:fplasma}
\end{eqnarray}
In the Earth's ionosphere, the electron number density is less than $10^6$~cm$^{-3}$, which implies that $f_{\rm plasma}^\oplus \lsim 10$~MHz.
At the favorable viewing condition, when the planet is on the near side of its star to the Earth, the maximum plasma frequency along the line of sight corresponds to that in the vicinity of the planet.
Therefore, combining Eq.~(\ref{eq:n}) and Eq.~(\ref{eq:fplasma}), we have
%%%
\begin{eqnarray}
\label{eq:fplasmalos} f_{\rm plasma}^{\rm los} & \sim & 8.98 {\rm~kHz} \times \left( \frac{\dot M_\star}{4 \pi a^2 m_p v} \times 1\rm~cm^3 \right)^{1/2} \\
\label{eq:fplasmalos_scaled} & = & 14.7{\rm~MHz} \times \left( \frac{a}{5~\mbox{AU}}\right)^{-1} \left(\frac{v}{30~\mbox{km/s}}  \right)^{-1/2} \notag \\
 & & \times \left( \frac{\dot M_\star}{10^{-8} M_{\odot }{\rm /yr}}\right)^{1/2} \, .
\end{eqnarray}
\memoYF{check numbers in above equation}
%%%
%where in Eq.~(\ref{eq:fplasmalos_scaled}) we have taken the solar mass-loss rate $\dot M_\odot$ to be $2.0 \times 10^{-14}~M_\odot$/yr and the typical solar wind velocity $v_{\odot }$ to be 400~km/sec.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[bp]
%   \plotonesc{radio_emission_Mdot1e-8_constRp_christensen.pdf}
%   \plotonesc{radio_emission_Mdot1e-5_constRp_christesnsen.pdf}
	\plotonesc{radio_emission_ChristensenAubert_multiplecriteria.pdf}
   \caption{Radio flux in unit of Jy from a planetary companion to a red giant with mass loss rate $10^{-8} M_{\odot }/\mbox{yr}$ (top) and an AGB star with mass loss rate $10^{-5} M_{\odot }/\mbox{yr}$ (bottom). The systems are located at 100 pc away. The doubly hashed regions show the parameter space where the planetary radio emission would not be observable at all because the maximum frequency of the emission (cyclotron frequency at the planetary surface, $f_{\rm cyc}$) is below the plasma frequency cut-off, $f_{\rm plasma}^{\rm los}$. The hashed regions with vertical lines show the parameter space where the frequencies of bulk radio emission, estimated as $\sim 0.1f_{\rm cyc}$, is below $f_{\rm plasma}^{\rm los}$, i.e., the large part of the emission would not be able to get out of the system. }
  \label{fig:radio}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

One of the key features that make planetary radio emission distinct is the almost complete polarization \citep[e.g.][]{dulk1994}.
...\memoYF{write here}

Another key is that auroral radio emission from planets is not a static source but it significantly varies in time. 
The potential factors that cause time variability of the auroral radio emission are  listed below.
